---
title: "Project proposal: Contaminated sediments in the Gulf of Maine"
author: "Joshua Harkness and Autumn Pauly,"
date: "October 2023"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(readxl)
```

## 1. Introduction
We will be looking at distributions and frequencies of contaminated sediments in the Gulf of Maine, using the Gulf of Maine Contaminated Sediments Database published by the U.S. Geological Survey in their Open-File Report 02-403.

There are seven data files in this database which we will use in some form, principally these include geographic information on stations (collection sites), textural sediments data, and contaminants data.

We are most interested in examining patterns in distribution and abundance of organic contaminants, which, in this dataset, are represented by polychlorinated biphenyls (PCBs), polycyclic aromatic hydrocarbons (PAHs), and organochlorine pesticides.  We may, however, examine inorganic contaminants, such as heavy metals, if we find that there is not as much organic contaminant data to be useful for our purposes here.


## 2. Data
```{r load-data, message=FALSE, warning=FALSE}
stations = read_excel("/cloud/project/data/STAT2002.xls", sheet = 2, skip = 3)
sediments = read_excel("/cloud/project/data/TXTR2002.xls", sheet = 2, skip = 3)
PCBs = read_excel("/cloud/project/data/PCBP2002.xls", sheet = 2, skip = 3)
PAHs = read_excel("/cloud/project/data/PAHS2002.xls", sheet = 2, skip = 3)
Organics = read_excel("/cloud/project/data/GENO2002.xls", sheet = 2, skip = 3)
Inorganics = read_excel("/cloud/project/data/INOR2002.xls", sheet = 2, skip = 3)
```

```{r}
stations %>%
  ggplot(aes(x = LONGITUDE, y = LATITUDE))+
  geom_point()+
  theme_bw()
```
Map of raw station data for Gulf of Maine, representing sediment sampling locations.  You can see the outline of the New England coast where points are dense, representing greater sampling.

```{r join-stations-organics}
StationOrganics = right_join(stations, Organics)
```
```{r}
BostonOrganics = filter(StationOrganics, GEN_LOC_NM %in% c("BOSTON INNER HARBOR","NORTHWEST BOSTON HARBOR","CENTRAL BOSTON HARBOR"))
```

```{r pcb-box1}
BostonOrganics %>%
  ggplot(aes(x = GEN_LOC_NM, y = PCB_T_UGG))+
  geom_boxplot()
```
```{r pcb-box1-zoomed}
BostonOrganics %>%
  ggplot(aes(x = GEN_LOC_NM, y = PCB_T_UGG))+
  geom_boxplot()+
  ylim(0,7.5)
```
Mean PCB abundance (ug/g) are less than 0.5 for Boston Inner Harbor, Central Boston Harbor, and Northwest Boston Harbor, but outliers are present as high as 3000ug/g.

## 3. Ethics review

## 4. Data analysis plan



